---
title: "learnr logerator"
author: "Peter Young"
date: "2025-10-19"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}

# note, this app is not for deployment

knitr::opts_chunk$set(echo = FALSE)

#library(httr2)
library(DBI)
library(RSQLite)  # you may export SQLite and connect locally
library(jsonlite)
library(purrr)
library(tidyverse)
library(purrr)
library(kableExtra)
library(httr2)
library(scales)

if (file.exists(".env")) {
  try(dotenv::load_dot_env(), silent = TRUE)
}
if (file.exists(".Renviron")) {
  try(dotenv::load_dot_env(".Renviron"), silent = TRUE)
} 

production_hosts <- Sys.getenv("SHINY_PRODUCTION_HOST", unset = "")
log_server_url <- Sys.getenv("CLOUDFLARE_LOGGER_D1_ENDPOINT", unset = "")
account_id <- Sys.getenv("CLOUDFLARE_ACCOUNT_ID") 
database_id <- Sys.getenv("CLOUDFLARE_LOGGER_D1_DB_ID") 
api_token <- Sys.getenv("CLOUDFLARE_LOGGER_D1_API_TOKEN")

cat(paste0("Account RENVIRON", Sys.getenv("READ_ACCT_RENVIRON")))
cat(paste0("Project RENVIRON", Sys.getenv("READ_PROJ_RENVIRON")))

# include local testing output?
exclude.local <- FALSE
start.date <- as.Date("2025-01-01") # can filter out older records

if(!nzchar(log_server_url)) {
  stop("CLOUDFLARE_LOGGER_D1_ENDPOINT environment variable not set")
}
if(!nzchar(account_id)) {
  stop("CLOUFLARE_ACCOUNT_ID environment variable not set")
}
if(!nzchar(database_id)) {
  stop("CLOUDFLARE_LOGGER_D1_DB_ID environment variable not set")
}
if(!nzchar(api_token)) {
  stop("CLOUDFLARE_LOGGER_D1_API_TOKEN environment variable not set")
}

query <- "select * from logs"

resp <- request(
  paste0(log_server_url, "/", account_id, "/d1/database/", database_id, "/query")
) |>
  req_headers(
    "Authorization" = paste("Bearer", api_token),
    "Content-Type"  = "application/json"
  ) |>
  req_body_json(list(sql = query)) |>
  req_perform()

# Parse
data <- resp_body_json(resp)

res <- data$result[[1]]$results
if (is.null(res) || !length(res)) {
  logs <- tibble()
} else {
  logs <- purrr::map_dfr(res, as_tibble)
}

```

```{r parse-logs, echo=FALSE}
## Logging outcomes

logs <- logs %>%
  mutate(info_parsed = map(info, ~ tryCatch(fromJSON(.x), error = function(e) NULL)))

# only include actual runs on production, and parse the json fields into sep cols
logs_clean <- logs %>% unnest_wider(info_parsed)

# earlier I was logging submissions instead of results, cleanup here
logs_clean <- logs_clean |> 
  mutate(event = if_else(event == "exercise_submitted", "exercise_result", event))

if(exclude.local == TRUE) {
  logs_clean <- logs_clean |> filter(hostname %in% production_hosts)
}
if(!is.na(start.date)) {
  logs_clean <- logs_clean |> filter(as.Date(ts) >= start.date)
}

# which tutorials actually have data?
found_tutorials <- logs_clean |> distinct(path) |> filter(!is.na(path)) |> 
  mutate(app = str_remove(path, "^(/tutorial-|/)"),
         app = str_remove(app, "/$"),
         app2 = str_replace_all(app, "-", " "),
         local_path = paste0(app, "/", app, ".Rmd"),
         local_path2 = paste0(app2, "/", app, ".Rmd")) |> 
  arrange(desc(path)) |>
  distinct(local_path, .keep_all = TRUE) # some hackery to deal with multiple deployments of same tutorial

# parse the total number of exercises and last exercise name from the source for those
# tutorials that were found
tutorial_func <-   function(path) {
    app = str_remove(path, "^(/tutorial-|/)")
    app = str_remove(app, "/$")
    app2 = str_replace_all(app, "-", " ")
    local_path = paste0(app, "/", app, ".Rmd")
    local_path2 = paste0(app2, "/", app, ".Rmd")
    
    #message(paste0("current directory is ", getwd()))
    suppressWarnings(tutorial_source <- tryCatch(readLines(local_path), error = function(e) NULL))
    if (is.null(tutorial_source)) {
      tutorial_source <- tryCatch(readLines(local_path2), error = function(e) NULL)
    }
    if (!is.null(tutorial_source)) {
      total_exercises <- sum(grepl("```\\{r,? .*, ?exercise=", tutorial_source))
      last_exercise <- tail(grep("```\\{r,? .*, ?exercise=", tutorial_source, value = TRUE), 1)
      last_exercise_name <- sub("```\\{r,? (.*), ?exercise=.*$", "\\1", last_exercise)
      return(data.frame(app = app, 
                        total_exercises = total_exercises, 
                        last_exercise_name = last_exercise_name))
    } else {
      return(data.frame(app = app, 
                        total_exercises = NA, 
                        last_exercise_name = NA))
  }
}
tutorial_detail_func <- function(path) {
  app = str_remove(path, "^(/tutorial-|/)")
  app = str_remove(app, "/$")
  app2 = str_replace_all(app, "-", " ")
  local_path = paste0(app, "/", app, ".Rmd")
  local_path2 = paste0(app2, "/", app, ".Rmd")
  
  suppressWarnings(tutorial_source <- tryCatch(readLines(local_path), error = function(e) NULL))
  if (is.null(tutorial_source)) {
    tutorial_source <- tryCatch(readLines(local_path2), error = function(e) NULL)
  }
  if (!is.null(tutorial_source)) {
    exercise_lines <- grep("```\\{r,? .*, ?exercise=", tutorial_source, value = TRUE)
    exercise_names <- sub("```\\{r,? (.*), ?exercise=.*$", "\\1", exercise_lines)
    return(data.frame(app = app, 
                      exercise_name = exercise_names, 
                      exercise_index = seq_along(exercise_names)))
  } else {
    return(data.frame(app = app, exercise_name = NA))
  }
}

# parse the tutorial structure from source code, details include ordered list of exercises
tutorial_info <- map_dfr(found_tutorials$path, tutorial_func)
tutorial_detail <- map_dfr(found_tutorials$path, tutorial_detail_func)

# get info on tutorial initiation
initiation <- tutorial_info |> left_join(logs_clean, join_by(app == app)) |> 
  filter(event == "session_start") |> 
  group_by(app) |> 
  summarise(initiations = n_distinct(session),
            last_init_date = as.Date(max(ts)),
            total_exercises = max(total_exercises),
            last_exercise_name = last(last_exercise_name),
            .groups = "drop")

# get info on tutorial completion (I'm not actually capturing this event, using last exercise
# as a proxy)
completion <- tutorial_info |> left_join(logs_clean, join_by(app == app, last_exercise_name == label)) |> 
  filter(event == "exercise_submitted") |> 
  group_by(app) |> 
  summarise(completions = n_distinct(session),
            .groups = "drop")

completion <- initiation |> left_join(completion, by = join_by(app)) |> 
  mutate(completions = if_else(is.na(completions), 0L, completions),
         completion_rate = percent(completions / initiations, 0.1))

exercise_detail <- 
  tutorial_detail |> 
  left_join(logs_clean |> filter(event %in% c("exercise_submitted", "exercise_result")), 
            join_by(app == app, exercise_name == label)) |> 
  group_by(app, exercise_name, exercise_index) |>
  summarise(sessions_attempting = n_distinct(session),
            correct_attempts = sum(correct == TRUE, na.rm = TRUE),
            total_attempts = sum(checked == TRUE, na.rm = TRUE),
            total_submissions = n(),
            last_submission_date = as.Date(max(ts)),
            prop_correct = percent(correct_attempts / total_attempts, 0.1),
            .groups = "drop") |> 
  arrange(app, exercise_index)

# for each session/app, compute proportion of exercises done and find last exercise
# completed

prop_exercise_unique_sessions <- 
  tutorial_info |> left_join(logs_clean, join_by(app == app)) |>
    filter(event == "exercise_submitted") |> 
    group_by(app, session) |> 
    summarise(exercises_done = n_distinct(label),
              total_exercises = first(total_exercises),
              last_exercise_date = as.Date(last(ts)), .groups = "drop") |> 
    mutate(prop_done = exercises_done / total_exercises)

# summarize the per-session data to see range of completion rates and when last 
# completed an exercise

prop_exercise_done <- 
  prop_exercise_unique_sessions |> group_by(app) |>
    summarise(min_prop_done = min(prop_done, na.rm = TRUE),
              avg_prop_done = mean(prop_done, na.rm = TRUE),
              max_prop_done = max(prop_done, na.rm = TRUE),
              last_ex_date = as.Date(max(last_exercise_date, na.rm=TRUE)),
              .groups = "drop") |> 
  mutate(across(ends_with("_done"), ~percent(.x, accuracy = 0.1)))

# fix up the percentage, now that we've used for computing ranges
sessions <-
  prop_exercise_unique_sessions |> 
  mutate(session = paste0(substr(session, 1, 6),  "...")) |>
  mutate(prop_done = percent(prop_done, accuracy = 0.1))

# create a nice time series for plotting
exercise_ts <-
  logs_clean |> mutate(week_start = lubridate::floor_date(as.Date(ts), "week")) |>
  group_by(week_start, app, event) |> 
  summarise(events = n(), .groups = "drop") # |> 
 # pivot_wider(names_from = event, values_from = events, values_fill = 0)

```

### Tutorial initiations and completions
```{r show-results-1}
completion |> 
  rename(
    "App" = app,
    "Starts" = initiations,
    "Last start" = last_init_date,
    "N. ex." = total_exercises,
    "Last ex." = last_exercise_name,
    "N. comp." = completions,
    "C. rate" = completion_rate
  ) |>
  kable(caption = "Initiations and completions")

```

### Tutorial progress
```{r show-results-2}
prop_exercise_done |>
    rename(
    "App" = app,
    "Min % done" = min_prop_done,
    "Avg % done" = avg_prop_done,
    "Max % done" = max_prop_done,
    "Last submit" = last_ex_date
  ) |>
    kable(caption = "Progress of exercises completed per session")
```

### Progress by session
```{r show-results-3}
sessions |>
    rename(
    "App" = app,
    "Session" = session,
    "Ex. done" = exercises_done,
    "N. ex." = total_exercises,
    "Last ex." = last_exercise_date,
    "% done" = prop_done
  ) |>
  kable(caption = "Session progress")
```

### Exercise attempts and correctness
```{r show-results-4}
exercise_detail |> 
  select(app, exercise_name, sessions_attempting, total_submissions, total_attempts, 
         prop_correct, last_submission_date) |>
    rename(
    "App" = app,
    "Ex. name" = exercise_name,
    "N. sessions" = sessions_attempting,
    "N. submissions" = total_submissions,
    "N. attempts" = total_attempts,
    "% correct" = prop_correct,
    "Last submit" = last_submission_date
  ) |>
  kable(caption = "Exercise attempts")
```

### Time series of tutorial intitiation and exercise completion
```{r show-results-5}
exercise_ts |> 
  ggplot(aes(x = week_start, y = events, color = event)) +
  geom_line() + geom_point() +
  facet_grid(~app) +
  theme_minimal() +
  scale_x_date(breaks = unique(exercise_ts$week_start)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```
